---
title: "Homework 5 Part 1"
author: "Jung Mee Park"
date: "11/8/2021"
output: html_document
---
# Part A
load dataset
```{r}
load("/Users/jungmeepark/Downloads/dataset.RData")
```
## section i. 
load packages
```{r include=FALSE}
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(arules)
library(factoextra)
library(cluster)
library(arulesViz)
```

1.	Compute the percentage of passengers that were children.
```{r}
titanic %>% 
  group_by(Age) %>% 
  summarise(Count=n()) %>% 
  mutate(Percentage=Count/sum(Count)*100)
```

2.	Compute the percentage of passengers that were women.
```{r}
titanic %>% 
  group_by(Sex) %>% 
  summarise(Count=n()) %>% 
  mutate(Percentage=Count/sum(Count)*100)
```

3.	Compute the percentage of passengers that were in first class.
```{r}
titanic %>% 
  group_by(Class) %>% 
  summarise(Count=n()) %>% 
  mutate(Percentage=Count/sum(Count)*100)
```

4.	Compute the percentage of people that survived.
```{r}
titanic %>% 
  group_by(Survived) %>% 
  summarise(Count=n()) %>% 
  mutate(Percentage=Count/sum(Count)*100)
```

5.	What percentage of children survived?
```{r}
options(digits = 2)
prop.table(table(titanic$Age, titanic$Survived),1)
```

6.	What percentage of women survived?
```{r}
prop.table(table(titanic$Sex, titanic$Survived),1)
```

7.	What percentage of first-class passengers survived?
```{r}
prop.table(table(titanic$Class, titanic$Survived),1)
```

8.	What percentage of third-class passengers survived?
```{r}
prop.table(table(titanic$Class, titanic$Survived),1) # This is the same table as above.
```

## section ii. 
9.	Create a function that subsets a given data frame by using sex, age, class and survival as arguments. This function should return a new data frame of passengers that satisfies a specified criteria.
```{r}
my_selector <- function(data=titanic, Class, Sex, Age, Survived){
  data[titanic$Class %in% Class & titanic$Sex %in% Sex & titanic$Age %in% Age & titanic$Survived %in% Survived, ]
}
```

10.	Using the previous function, create a function that calculates the percentages of people who lived or died:
```{r}
pct_survived <- function(data=titanic, Class, Sex, Age){
  num <- nrow(my_selector(Class = Class, Sex = Sex, Age = Age, Survived = "Yes"))
  den <- num + nrow(my_selector(Class = Class, Sex = Sex, Age = Age, Survived = "No"))
  pct <- (num/den)*100
  return(pct)
}
```

a.	Compare survival rates between men and boys in third-class fares.
```{r}
pct_survived(Class = "3rd", Sex = "Male", Age = "Adult")
```
```{r}
pct_survived(Class = "3rd", Sex = "Male", Age = "Child")
```

- Men in 3rd class pretty much all perished. Men in the 3rd class had a low survival rate of 16%. Boys from the 3rd class only had a slightly higher survival rate at 27%.

b.	Compare survival rates between women and girls in first-class fares.
```{r}
pct_survived(Class = "1st", Sex = "Female", Age = "Adult")
```
for girls
```{r}
pct_survived(Class = "1st", Sex = "Female", Age = "Child")
```
- Women and girls in the first class pretty much all survived. 100% of the girls survived. 97% of the women survived. 

## section iii. 
11.	Calculate some association rules using apriori() command. Please focus on the passengers who survived only: 
```{r}
titanic_rules <- apriori(titanic, parameter = list(minlen=3,supp=0.002,conf=0.2), 
                         control = list(verbose=F), 
                         appearance = list(default="lhs", rhs=("Survived=Yes")))
inspect(titanic_rules)
```
THe output above shows that children of the 1st and 2nd class have the same survival rate (Zhao 2013, 99).

12.	Please visualize the results (use the arulesViz). Add captions to each of the figures you generate. Explain the meaning of the axes.
```{r message=FALSE}
plot(titanic_rules,method="scatterplot",measure=c("support","confidence"),shading=c("lift"))
```

Scatter plot of the association rules. 

```{r, fig.cap = "Left-hand side (LHS) shows several combinations of factors. Right-hand side (RHS) in this case represents survival." }
plot(titanic_rules,method="grouped")
```

grouped by confidence
```{r message=FALSE, fig.cap="Ordered by survival rate, children of the 1st and 2nd classes all survived and first class women nearly all survived."}
plot(titanic_rules,method="grouped",measure=c("confidence")) 
```
The post likely to survive were children of 1st and 2nd classes. 

graph of association rules
```{r warning=FALSE, fig.cap="Children and women of the 1st and 2nd classes survived moreso than men of the corresponding classes."}
plot(titanic_rules, method="graph")
```

graph of association rules with control type items
```{r warning=FALSE,message=FALSE, fig.cap="Since this is a graph of all those who survived, we can see that the crew, mostly made up of men, mostly did not survive."}
plot(titanic_rules, method="graph", control=list(type="items"))
```

13.	Sort the list of previously created rules. Use a lift > 3.0 to find the best rules. Briefly explain the best set of rules.
```{r}
rules.subset2 <- subset(titanic_rules, lift > 3)
inspect(rules.subset2)
```

14.	Using the function in Q10, estimate the percentage of survival for each of the best rules selected in Q13. Refer to HW4 to re-assess your previous answers (+2 points).
```{r}
pct_survived(Class = "1st", Sex = "Male", Age = "Child")
```
```{r}
pct_survived(Class = "2nd", Sex = "Male", Age = "Child")
```
```{r}
pct_survived(Class = "1st", Sex = "Female", Age = "Child")
```
```{r}
pct_survived(Class = "2nd", Sex = "Female", Age = "Child")
```
```{r}
pct_survived(Class = "1st", Sex = "Female", Age = "Adult")
```
```{r}
pct_survived(Class = "2nd", Sex = "Female", Age = "Adult")
```

### Reference
Zhao, Y. (2013). R and Data Mining: Examples and Case Studies. Elsevier Science & Technology.
