---
title: "HW_4_Park"
author: "Jung Mee Park"
date: "11/1/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---
# Part 1.
Load a basic Titanic dataset. Use the following code:
```{r include=FALSE}
library(datasets)
library(psych)
library(GGally)
library(Amelia)
library(ggplot2)
library(dplyr)
library(plotly)
library(corrplot)
```

```{r}
data(Titanic)
```

1.	What is the class of object “Titanic”? What variables does it have?  Explain. (1 pt)
* There are categorical variables, which are classified as factors or characters, in R. We see variables about what class, the passengar was in - 1st, 2nd, or 3rd, whether the person was an adult or child, sex, and whether the passengar survived or not. 

2.	Run this code: ‘plot(Titanic)’. What kind of plot did you get? Why was the data plotted this way? What is your interpretation of this graph? (1 pt)
```{r}
plot(Titanic)
```

* The plot allows you to see survival rates based on different variables such as sex, class, and age. From what I see, women survived at a higher rate than men, especially if they were in first class. First class passengers survived more than the other classes. 

3.	Transform the data into a data.frame. We’ll be using this object to create some visuals in the next questions. (1 pt)
```{r}
titanic <- as.data.frame(Titanic)
```

4.	(ggplot2) Now, please create a canvas for the figure. Show only the axes (i.e. don’t include any geometry). Plot “class” (x axis) and “freq” (y axis). (1 pt)

```{r}
ggplot(titanic, aes(Class, Freq)) 
```

5.	 (ggplot2) Display the data as a boxplot. Can you think of a better (or similar) way to display the visual patterns between “class” and “freq”? (2 pts) Explain.
```{r}
ggplot(titanic, aes(Class, Freq)) +
  geom_boxplot()
```

6.	(ggplot2) Now, include the raw data on top of the boxplot. Use the geom_jitter() and list one additional alternative for plotting the raw data. Please make sure that outliers are not represented twice (e.g. plotted under the boxplot and in the scatterplot layers). (1 pt)
```{r}
ggplot(titanic, aes(Class, Freq)) +
  geom_boxplot() +
  geom_jitter()
```

7.	(ggplot2) In geom_jitter, set ‘width=0.25’ and ‘alpha=0.5’. Explain what these arguments do. (1 pt)
```{r}
ggplot(titanic, aes(Class, Freq)) +
  geom_boxplot() +
  geom_jitter(width = 0.25, alpha=0.5)
```

8.	(ggplot2) Plot “freq” by “class” using histograms. Make only a single plot. Change the colors. (1 pt)
```{r}
library(RColorBrewer)
ggplot(titanic, aes(Freq, fill = Class)) + 
  geom_histogram(binwidth = 35) +
  scale_fill_brewer(palette = "Set3") 
```

9.	Perform steps 4–6, and 8 using functions from base R. (5 pts)
```{r}
boxplot(Freq ~ Class, method = "jitter", pch = 19,
        col = 4, data=titanic)
```

10.	Perform steps 4–6, and 8 using plotly from R. (5 pts)
```{r}
p1 <- plot_ly(titanic, x = ~Freq, type = "histogram") 
p1 <- p1 %>%  layout(barmode = "overlay")
p1 
```

# Part 2.
## Titanic dataset from Kaggle

1.	What was the survival frequency on Titanic? Examine this question from the following variables: gender, class, age, and ticket fare. 
Load in data
```{r}
setwd("~/Documents/INFO 523 fall 2021/Assignments/titanic")
train_titanic <- read.csv("train.csv")
```

Transform select variables into factors
```{r}
train_titanic$Survived  <- recode_factor(train_titanic$Survived , "1" = "Survived", "0" = "Died")
train_titanic$Pclass <- factor(train_titanic$Pclass, levels = c("3","2","1"), 
                                  ordered = TRUE)
train_titanic$PassengerId <- as.factor(train_titanic$PassengerId)
train_titanic$Cabin <- as.factor(train_titanic$Cabin)
```
Prepare missing variables
```{r}
train_titanic[train_titanic == ""] <- NA                     # Replace blank by NA
```
Check for missing values
```{r}
missmap(train_titanic, main = "Missing values vs observed") # age is missing in some cases
```

Survival count by sex
```{r}
gender_plot <- ggplot(train_titanic, aes(x=Survived, fill=Sex)) +
  geom_bar(position=position_dodge()) +
  geom_text(stat = 'count',
            aes(label=stat(count)),
            position = position_dodge(width = 1), vjust=-0.5) + 
            xlab("outcome") 
gender_plot <- gender_plot + labs(title = "Survival count by sex", 
                                      subtitle = "training dataset") 
gender_plot
```
Survival by class
```{r}
class_graph <- ggplot(train_titanic, aes(x=Survived, fill=Pclass)) + 
  geom_bar(position = position_dodge()) + 
  geom_text(stat = 'count',
            aes(label=stat(count)),
            position = position_dodge(width = 1), vjust=-0.5) + 
  xlab("outcome") 


class_graph <- class_graph + labs(title = "Survival count by class",
                                    # subtitle = "training dataset", 
                                    fill = "Class")
print(class_graph)
```
Survival by age
```{r}
train_titanic$Discretized.age <- cut(train_titanic$Age, c(0,10,20,30,40,50,60,70,80,100))

age_bar <- ggplot(train_titanic, aes(x = Discretized.age, fill=Survived)) +
  geom_bar(position = position_dodge()) +
  geom_text(stat='count', aes(label=stat(count)), position = position_dodge(width=1), vjust=-0.5)+
  theme_classic() + 
  xlab("outcome") 

age_bar <- age_bar + labs(title = "Survival count by age",
                                  subtitle = "training dataset", 
                                  fill = "Survived")
print(age_bar)

```
Survival by fare
```{r}

violin_first <- ggplot(train_titanic, aes(x = Survived, y = Fare, fill=Sex)) +
  geom_violin() +
  ylab("cost of fare") 
violin_first <- violin_first +  labs(title = "Violin plot of Survival by fare and gender", 
                                     subtitle = "training dataset") 
print(violin_first)
```


If possible, compare groups using the appropriate statistical tests (e.g. t-student vs Wilcoxon-Mann-Whitney U).


2.	What was the order of priority (i.e. gender and age) to rescue passengers of Titanic?
```{r}
full_titantic_model <- glm(Survived ~ Sex+Pclass+Age+SibSp+Parch+Fare,
                     data= train_titanic,       
                     family="binomial")         

summary(full_titantic_model)
```
calculate predicted probabilities in a results table
```{r}
predicted_probabilities <- predict(full_titantic_model,              
                                   newdata=train_titanic,      
                                   type="response")   
### Convert to 0, 1 preds
class_preds <- ifelse(predicted_probabilities >= 0.5, 1, 0)  

result_table <- table(class_preds,             
                      train_titanic$Survived)  

result_table
```


3.	What were the characteristics of the passengers that did not survive? Approach this question from the angle of class, age, and fare distribution.
```{r}
violin_two <- ggplot(train_titanic, aes(x = Survived, y = Fare)) +
  geom_violin() +
  ylab("cost of fare") +
  stat_summary(fun = "mean", geom = "point", shape = 8, size = 3, color = "midnightblue") 

violin_two <- violin_two +  labs(title = "Violin plot of Survival by fare", 
                                     subtitle = "training dataset") 
print(violin_two)
```
looking at sex and age 
```{r}
violin_age <- ggplot(train_titanic, aes(x = Survived, y = Age, fill = Sex)) +
  geom_violin() +
  # stat_summary(fun = "mean", geom = "point", shape = 8, size = 3, color = "midnightblue") +
  ylab("Age") 
violin_age <- violin_age +  labs(title = "Violin plot of Survival by age and sex", 
                                 subtitle = "training dataset") 
print(violin_age)
```
* Men who paid lower fares tended not to survive. Also men under 40 did not survive. Men who died had a younger mean age at 

4.	Do you notice any interesting patterns in survival frequencies based on the place where passengers boarded? (e.g. where did most passengers board? Is there any difference in survival for passengers boarding from different places?)
```{r}
embark_graph <- ggplot(train_titanic, aes(x=Survived, fill=Embarked)) + 
  geom_bar(position = position_dodge()) + 
  geom_text(stat = 'count',
            aes(label=stat(count)),
            position = position_dodge(width = 1), vjust=-0.5) + 
  xlab("outcome") 

embark_graph <- embark_graph + labs(title = "Survival count by port of embarkment",
                                  subtitle = "training dataset", 
                                  fill = "Port")
print(embark_graph)
```
Port of embarkment distribution 
```{r}
table(train_titanic$Embarked) # most people embarked in Southampton
table(train_titanic$Embarked, train_titanic$Survived)
```
Most people boarded in Southampton, so there may habe been anecdotal evidence that people who boarded elsewhere survived more often. 55% of those who embarked in Cherbourg survived and only 34% of those who boarded in Southampton survived. Queenstown had a survival percent of 39%. 

5.	Was it a good idea to travel in big groups with parents/children or was it better to travel in small groups with siblings/spouses?
```{r}

```

